---
title: "Tidying and Transforming Data"
author: "MD Asadul Islam"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(tidyr)
library(dplyr)
library(ggplot2)
```

### Read data from the CSV file

```{r code-chunk-label}
# Read the csv file
data <- read.csv("C:\\Users\\HP\\Documents\\arrival delays for two airlines.csv")
print(data)

```
The above section reads data from a CSV file and inspects it for any inconsistencies
### Tidy the data using tidyr and diplyr

```{R code chunk for tidying data}
tidy_airline_data <- data %>%
 pivot_longer(cols = c(`Los.Angeles`, Phoenix, `San.Diego`, `San.Francisco`, Seattle), 
               names_to = "Destination", 
               values_to = "Count") %>%
  arrange(Airline, Status, Destination)

print(tidy_airline_data)
```
In this section the data was tidied using the `tidyr` and `dplyr` packages in R. The `pivot_longer` function was used to reshape the data into a long format, making it easier to analyze and visualize

### Analysis to Compare the Arrival Delays

```{r code chunk for analysis}
# Summarize the on-time and delayed counts by airline
summary_delays <- tidy_airline_data %>%
  group_by(Airline, Status) %>%
  summarize(Total_Count = sum(Count))
print(summary_delays)

# Plot the delays
ggplot(tidy_airline_data, aes(x = Destination, y = Count, fill = Status)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ Airline) +
  theme_minimal() +
  labs(title = "Comparison of Arrival Delays by Airline",
       x = "Destination",
       y = "Count",
       fill = "Status")

```

The above section has the summary statistics. The summary statistics for the on-time and delayed counts were calculated for each airline. The total counts were determined for both statuses. Additionally, a bar plot was created to visually compare the arrival delays for the two airlines across different destinations.

### Conclusion

Based on the analysis, it appears that Am West has a significantly higher count of on-time arrivals compared to Alaska, especially for the Phoenix destination. However, Am West also has a higher count of delays compared to Alaska. The visualizations provide a clear comparison of the performance of the two airlines across different destinations for the two statuses.

...

